### 2015-03-16 Kokkos_PG_Introduction.tex

\chapter{Introduction}\label{C:intro}

The field of High Performance Computing (HPC) is entering a new era.
Since the mid 1990's, HPC applications have had great success 
with distributed-memory parallel computing using the Message-Passing Interface (MPI).
However, after two decades of MPI dominance, 
increasing core counts on standard processors and the emergence of new throughput-oriented processors, such as GPUs, make a new approach necessary.
MPI will not necessarily go away,
but application developers will need to expose finer-grained shared-memory parallelism
\emph{as well as} coarser-grained MPI-level parallelism,
in order to get their money's worth on new architectures.

Developers have many shared-memory parallel programming model options today.
Those include OpenMP, OpenACC, CUDA, TBB, Cilk, OpenCL, and C++AMP, among others.
Some of these programming models are vendor specific, while others have varying degrees of support on different hardware. 
Most of the models lack a concept to handle complex memory hierarchies, 
which in addition to threading is one of the defining upcoming trends in HPC.

The Kokkos programming model addresses the increasing complexity and variety of hardware architectures
through a rigorous introduction of parallel execution and memory handling abstractions.
These abstraction layers isolate a software developer from hardware details and allow the implementation of performance-portable applications.

This Guide covers both the programming model, and its implementation as an embedded C++ library.
It describes the abstraction concepts employed in Kokkos and their exposure as a C++ interface.
We intend it as an introduction for new users as well as a reference for application developers already using Kokkos.
We have also supplemented the Guide with tutorial examples in the \verb!example! subdirectory of Kokkos' source code.