### 2015-03-16 Kokkos_PG_MachineModel.tex

\chapter{Machine Model}

The abstract machine model of Kokkos allows for very complex heterogenous node architectures containing N execution spaces and M memory spaces.