### 2015-03-16 Kokkos_PG_Subviews.tex

\chapter{Subviews}\label{C:Subviews}

After reading this chapter, you should understand the following:
\begin{itemize}
\item A \emph{slice} of a multidimensional array behaves as an array,
  and is a view of a structured subset of that array
\item A \emph{subview} is a slice of an existing Kokkos View
\item A subview has the same reference count as their parent View
\item Use C++11 \texttt{auto} to let Kokkos pick the subview's type
\end{itemize}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{A subview is a slice of a View}\label{S:Subviews:Slice}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

A \emph{slice} of a multidimensional array behaves as an array, and is
a view of a structured subset of the original array.  ``Behaves as an
array'' means that the slice has the same syntax as an array should;
one can access its entries using array indexing notation.  ``View''
means that the slice and the original array point to the same data.
That is, the slice sees changes to the original array, and vice versa.
``Structured subset'' means a cross product of indices along each
dimension, as for example a plane or face of a cube.  If the original
array has dimensions $(N_0, N_1, \dots, N_{k-1})$, then a slice views
all entries whose indices are $(\alpha_0, \alpha_1, \dots,
\alpha_{k-1})$, where $\alpha_j$ is a subset of $[0, 1, \dots,
N_j-1]$.

Array slices are handy for encapsulation.  A slice looks and acts like
an array, so you can pass it into functions that expect an array.  For
example, you can write a function for processing boundaries (as
slices) of a structured grid, without needing to tell that function
properties of the entire grid.

Programming languages like Fortran 90, Matlab, and Python have a
special ``colon'' notation for representing slices.  For example, if
\texttt{A} is an $M \times N$ array, then
\begin{itemize}
\item \texttt{A(:, :)} represents the whole array,
\item \texttt{A(:, 3)} represents the fourth column (if the language
  has zero-based indices, or the third column, if the language has
  one-based indices),
\item \texttt{A(4, :)} represents the fifth row, 
\item \texttt{A(2:4, 3:7)} represent the sub-array of rows 3-4 and
  columns 4--7 (different languages differ on whether the ranges are
  inclusive or exclusive of the last index -- Kokkos, like Python, is
  exclusive), and
\item \texttt{A(3, 4)} represents a ``zero-dimensional'' slice which
  views the entry in the fourth row and fifth column of the matrix.
\end{itemize}
These languages may have more elaborate notation for expressing sets
of indices other than contiguous ranges.  These may include
``strided'' subsets of indices, like \texttt{3:2:9} = $\{ 3, 5, 7, 9
\}$, or even arbitrary sets of indices.

In Kokkos, a \emph{subview} is a slice of a View.  A subview has the
same reference count as its parent View, so the parent View won't be
deallocated before all subviews go away.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{How to take a subview}\label{S:Subviews:How}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

To take a subview of a View, use the \texttt{subview} function.  This
function is overloaded for all different kinds of Views and index
ranges.  For example, the following is equivalent to the above example
\texttt{A(2:4, 3:7)}:

\begin{lstlisting}
const size_t N0 = ...;
const size_t N1 = ...;
View<double**> A ("A", N0, N1);

auto A_sub = 
  subview (A, std::make_pair (2, 4), std::make_pair (3, 7));
\end{lstlisting}

The Kokkos equivalent of a contiguous index range \texttt{3:7} is
\\ \lstinline!std::pair<size_t, size_t>(3, 7)!.  The Kokkos equivalent of
\texttt{:} (a colon by itself; the whole index range for that
dimension) is \texttt{ALL()} (an instance of the \texttt{Kokkos::ALL}
class).  Kokkos does not currently have equivalents of the strided or
arbitrary index sets.

Every subview is also a View.  This means that you may take a subview
of a subview.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{C++11 type deduction}\label{SS:Subviews:How:auto}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Note the use of the C++11 keyword \texttt{auto} in the above example.
A subview may have a different type than its parent View.  For
instance, if \lstinline!A! has \lstinline!LayoutRight!,
\lstinline!A_sub!  has \lstinline!LayoutStride!.  A subview of an
entire row of \lstinline!A! has \lstinline!LayoutRight! as well, but a
subview of an entire column of \lstinline!A! has
\lstinline!LayoutStride!.  If you assign the result of the
\texttt{subview} function to the wrong type, Kokkos will emit a
compile-time error.  The best way to avoid this is to use the C++11
\texttt{auto} keyword to let the compiler deduce the correct type for
you.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Dimension of a subview}\label{SS:Subviews:How:dim}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Suppose that a View has $k$ dimensions.  Then, when calling
\texttt{subview} on that View, you must pass in $k$ arguments.  Every
argument that is a single index -- that is, not a pair or
\lstinline!ALL()! -- reduces the dimension of the resulting View by 1.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Degenerate Views}\label{SS:Subviews:How:degen}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Given a View with $k$ dimensions, we call that View \emph{degenerate}
if any of those dimensions is zero.  Degenerate Views are useful for
keeping code simple, by avoiding special cases.  For example, in an
MPI (Message-Passing Interface) distributed-memory parallel code that
uses Kokkos to represent local (per-process) data structures,