### 2015-03-16 Kokkos_PG_Views.tex

\chapter{View: Multidimensional array}

After reading this chapter, you should understand the following:
\begin{itemize}
\item A Kokkos View is an array of zero or more dimensions
\item How to use View's first template parameter to specify the type
  of entries, the number of dimensions, and whether the dimensions are
  determined at run time or compile time
\item Kokkos handles array deallocation automatically
\item Kokkos chooses array layout at compile time for best overall
  performance, as a function of the computer architecture
\item How to set optional template parameters of View for low-level
  control of execution space, layout, and how Kokkos accesses array
  entries
\end{itemize}
In all code examples in this chapter, we assume that all classes in
the \lstinline!Kokkos! namespace have been imported into the working
namespace.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Why Kokkos needs multidimensional arrays}\label{S:Views:Why}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Many scientific and engineering codes spend a lot of time computing
with arrays of data.  Programmers invest a lot of effort making array
computations as fast as possible.  This effort is often intimately
bound to details of the computer architecture, run-time environment,
language, and programming model.  For example, optimal array layout
may differ based on the architecture, with a large integer factor
penalty if wrong.  Low-level issues like pointer alignment, array
layout, indexing overhead, and initialization all affect performance.
This is true even with sequential codes.  Thread parallelism adds even
more pitfalls, like first-touch allocation and false sharing.

For best performance, coders need to tie details of how they manage
arrays, to details of how parallel code accesses and manages those
arrays.  Programmers who write architecture-specific code then need to
mix low-level features of the architecture and programming model into
high-level algorithms.  This makes it hard to port codes between
architectures, or to write a code that still performs well as
architectures evolve.

Kokkos aims to relieve some of this burden, by optimizing array
management and access for the specific architecture.  Tying arrays to
shared-memory parallelism lets Kokkos optimize the former to the
latter.  For example, Kokkos can easily do first-touch allocation,
because it controls threads that it can use to initialize arrays.
Kokkos' architecture awareness lets it pick optimal layout and pad
allocations for good alignment.  Expert coders can also use Kokkos to
access low-level or more architecture-specific optimizations in a more
user-friendly way.  For instance, Kokkos makes it easy to experiment
with different array layouts.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Creating and using a View}\label{S:Views:CreateUse}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Constructing a View}

A View is an array of zero or more dimensions.  Programmers set both
the type of entries, and the number of dimensions, at compile time, as
part of the type of the View.  For example, the following specifies
and allocates a View with four dimensions, whose entries have type
\texttt{int}:

\begin{lstlisting}
const size_t N0 = ...;
const size_t N1 = ...;
const size_t N2 = ...;
const size_t N3 = ...;
View<int****> a ("some label", N0, N1, N2, N3);
\end{lstlisting}

The string argument is a label, which Kokkos uses for debugging.
Different Views may have the same label.  The ellipses indicate some
integer dimensions specified at run time.  Users may also set some
dimensions at compile time.  For example, the following View has two
dimensions.  The first one (represented by the asterisk) is a run-time
dimension, and the second (represented by [3]) is a compile-time
dimension.  Thus, the View is an N by 3 array of type double, where N
is specified at run time in the View's constructor.

\begin{lstlisting}
const size_t N = ...;
View<double*[3]> b ("another label", N);
\end{lstlisting}

Views may have up to (at least) 8 dimensions, and any number of these
may be run-time or compile-time.  The only limitation is that the
run-time dimensions (if any) must appear first, followed by all the
compile-time dimensions (if any).  For example, the following are
valid three-dimensional View types:
\begin{itemize}
\item \lstinline!View<int***>!  (3 run-time dimensions)
\item \lstinline!View<int**[8]>!  (2 run-time, 1 compile-time)
\item \lstinline!View<int*[3][8]>!  (1 run-time, 2 compile-time)
\item \lstinline!View<int[4][3][8]>!  (3 compile-time)
\end{itemize}
and the following are \emph{not} valid three-dimensional View types:
\begin{itemize}
\item \lstinline!View<int[4]**>!
\item \lstinline!View<int[4][3]*>!
\item \lstinline!View<int[4]*[8]>!
\item \lstinline!View<int*[3]*>!
\item \lstinline!View<int*[3][8]>!
\end{itemize}
This limitation comes from the implementation of View using C++
templates.  View's first template parameter must be a valid C++ type.

\subsection{What types of data may a View contain?}

C++ lets users construct data types that may ``look like'' numbers in
terms of syntax, but do arbitrarily complicated things inside.  Some
of those things may not be thread safe, like unprotected updates to
global state.  Others may perform badly in parallel, like fine-grained
dynamic memory allocation.  As a result, Kokkos restricts the set of
data types \texttt{T} which may be the entries of a View.

Users may always construct a View whose entries are 
\begin{itemize}
\item built-in data types (``plain old data''), like \texttt{int} or
  \texttt{double}, or
\item structs of built-in data types.
\end{itemize}
The latter specifically forbids nontrivial destructors, constructors
with dynamic memory allocation, or virtual methods of any kind.  In
particular, entries must have size determined at compile time by the
\texttt{sizeof} operator.\footnote{Kokkos has a customization facility
  for users to add support for types with size determined at run time,
  but this is for expert users only.  For examples, see the Stokhos
  software package in the Trilinos project.}

\subsection{Accessing entries (indexing)}

You may access an entry of a View using parentheses enclosing a
comma-delimited list of integer indices.  This looks just like a
Fortran multidimensional array access.  For example:
\begin{lstlisting}
const size_t N = ...;
View<double*[3][4]> a ("some label", N);
parallel_for (N, [=] (const ptrdiff_t i) {
    const size_t j = ...;
    const size_t k = ...;
    const double a_ijk = a(i,j,k);
    /* rest of the loop body */
  });
\end{lstlisting}
Note how in the above example, we only access the View's entries in a
parallel loop body.  In general, you may only access a View's entries
in an execution space which is allowed to access that View's memory
space.  For example, if the default memory space is \lstinline!Cuda!
(CUDA device memory), you may not access it in host code, unless CUDA
UVM is enabled.  Furthermore, access costs (e.g., latency and
bandwidth) may vary, depending on the View's ``native'' memory and
execution spaces, and the execution space from which you access it.
CUDA UVM may work, but it may also be slow, depending on your access
pattern and performance requirements.  Thus, best practice is to
access the View only in a Kokkos parallel for, reduce, or scan, using
the same execution space as the View.  This also ensures that access
to the View's entries respect first-touch allocation.  The first
(leftmost) dimension of the View is the \emph{parallel dimension},
over which it is most efficient to do parallel array access.

\subsection{Memory management and copying}

\subsubsection{Reference counting}

Kokkos automatically manages deallocation of Views, through a
reference-counting mechanism.  Otherwise, Views behave like raw
pointers.  Copying or assigning a View does a shallow copy, and
changes the reference count.  (The View copied has its reference count
incremented, and the assigned-to View has its reference count
decremented.)  A View's destructor (called when the View falls out of
scope or during a stack unwind due to an exception) decrements the
reference count.  Once the reference count reaches zero, Kokkos may
deallocate the View.

For example, the following code allocates two Views, then assigns one
to the other.  That assignment may deallocate the first View, since it
reduces its reference count to zero.  It then increases the reference
count of the second View, since now both Views point to it.
\begin{lstlisting}
View<int*> a ("a", 10);
View<int*> b ("b", 10);
a = b; // assignment does shallow copy
\end{lstlisting}

For efficiency, View allocation and reference counting turn off inside
of Kokkos' parallel for, reduce, and scan operations.  This affects
what you can do with Views inside of Kokkos' parallel operations.

\subsubsection{Deep copy}

If you want to make a deep copy of a View's contents, you must do so
explicitly, using the \lstinline!deep_copy! function.  For example:
\begin{lstlisting}
View<int*> a ("a", 10);
View<int*> b ("b", 10);
deep_copy (a, b); // copy contents of b into a
\end{lstlisting}

\subsection{Where and how is the View allocated?}

\subsubsection{Execution and memory spaces}

Views are allocated by default, in the default execution space's
default memory space.  You may access the View's execution space via
its \lstinline!execution_space! typedef, and its memory space via its
\lstinline!memory_space! typedef.  You may also specify the execution space
explicitly as a template parameter.  For example, the following
allocates a View in CUDA device memory:
\begin{lstlisting}
View<int*, Cuda> a ("a", 100000);
\end{lstlisting}
and the following allocates a View in ``host'' memory, using OpenMP
for first-touch allocation:
\begin{lstlisting}
View<int*, OpenMP> a ("a", 100000);
\end{lstlisting}

\subsubsection{Initialization}

A View's entries are initialized to zero by default.  Initialization
happens in parallel, for first-touch allocation over the first
(leftmost) dimension of the View.

You may allocate a View without initializing.  For example:
\begin{lstlisting}
View<int*> x (ViewAllocateWithoutInitializing (label), 100000);
\end{lstlisting}

\subsection{How do I get the raw pointer?}

We discourage access to a View's ``raw'' pointer.  This circumvents
reference counting.  That is, the memory may be deallocated once the
View's reference count goes to zero, so holding on to a raw pointer
may result in invalid memory access.  Furthermore, it may not even be
possible to access the View's memory from a given execution space.
For example, a View in the \lstinline!Cuda! space points to CUDA device
memory.  Nevertheless, sometimes you really need access to the
pointer.  For such cases, we provide the \lstinline!ptr_on_device()!
method.  For example:
\begin{lstlisting}
// Legacy function that takes a raw pointer.
extern void legacyFunction (double* x_raw, const size_t len);

// Your function that takes a View.
void myFunction (const View<double*>& x) {
  // DON'T DO THIS UNLESS YOU MUST
  double* a_raw = a.ptr_on_device ();
  const size_t N = x.dimension_0 ();
  legacyFunction (a_raw, N);
}
\end{lstlisting}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Layout}\label{S:Views:Layout}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Strides and dimensions}\label{SS:Views:Layout:Strides}

\emph{Layout} refers to the mapping from a logical multidimensional
index $(i, j, k, \dots)$ to a physical memory offset.  Different
programming languages may have different layout conventions.  For
example, Fortran uses \emph{column-major} or ``left'' layout, where
consecutive entries in the same column of a 2-D array are contiguous
in memory.  Kokkos calls this \lstinline!LayoutLeft!.  C, C++, and Java use
\emph{row-major} or ``right'' layout, where consecutive entries in the
same row of a 2-D array are contiguous in memory.  Kokkos calls this
\lstinline!LayoutRight!.

The generalization of both left and right layouts is ``strided.''  For
a strided layout, each dimension has a \emph{stride}.  The stride for
that dimension determines how far apart in memory two array entries
are, whose indices in that dimension differ only by one, and whose
other indices are all the same.  For example, with a 3-D strided view
with strides $(s_1, s_2, s_3)$, entries $(i, j, k)$ and $(i, j+1, k)$
are $s_2$ entries (not bytes) apart in memory.  Kokkos calls this
\lstinline!LayoutStride!.

Strides may differ from dimensions.  For example, Kokkos reserves the
right to pad each dimension for cache or vector alignment.  You may
access all the dimensions of a View using the \lstinline!dimension! method,
and all the strides using the \lstinline!stride! method.  Both take a raw
integer array, and only fill in as many entries as the number of
dimensions.  For example:
\begin{lstlisting}
const size_t N0 = ...;
const size_t N1 = ...;
const size_t N2 = ...;
View<int***> a ("a", N0, N1, N2);
size_t dims[3]
a.dimension (dims); // fill 'dims' with dimensions
size_t strides[3]
a.strides (dims); // fill 'strides' with strides
\end{lstlisting}
You may also refer to specific dimensions:
\begin{lstlisting}
const size_t n0 = a.dimension_0 ();
const size_t n2 = a.dimension_2 ();
\end{lstlisting}

Users of the BLAS and LAPACK libraries may be familiar with the ideas
of layout and stride.  These libraries only accept matrices in
column-major format.  The stride between consecutive entries in the
same column is 1, and the stride between consecutive entries in the
same row is \lstinline!LDA! (``leading dimension of the matrix A'').  The
number of rows may be less than \lstinline!LDA!, but may not be greater.

\subsection{Other layouts}\label{SS:Views::Layout::Other}

Other layouts are possible.  For example, Kokkos has a ``tiled''
layout, where a tile's entries are stored contiguously (in either row-
or column-major order) and tiles have compile-time dimensions.  One
may also use Kokkos to implement Morton ordering or variants thereof.

\subsection{Default layout depends on execution space}\label{SS:Views:Layout:Default}

Kokkos selects a View's default layout for optimal parallel access
over the leftmost dimension, based on its execution space.  For
example, \lstinline!View<int**, Cuda>! has \lstinline!LayoutLeft!, so that
consecutive threads in the same warp access consecutive entries in
memory.  This \emph{coalesced access} gives the code better memory
bandwidth.  In contrast, \lstinline!View<int**, OpenMP>! has
\lstinline!LayoutRight!, so that a single thread accesses contiguous
entries of the array.  This avoids wasting cache lines and helps
prevent false sharing of a cache line between threads.

\subsection{Explicitly specifying layout}\label{SS:Views:Layout:Explicit}

We prefer that users let Kokkos determine a View's layout, based on
its execution space.  However, sometimes you really need to specify
the layout.  For example, the BLAS and LAPACK libraries only accept
column-major arrays.  If you want to give a View to the BLAS or
LAPACK, that View must be \lstinline!LayoutLeft!.  You may specify the
layout as a template parameter of View.  For example:
\begin{lstlisting}
const size_t N0 = ...;
const size_t N1 = ...;
View<double**, LayoutLeft> A ("A", N0, N1);

// Get 'LDA' for BLAS / LAPACK
int strides[2]; // any integer type works in stride()
A.stride (strides);
const int LDA = strides[1];
\end{lstlisting}
You may ask a View for its layout via its \lstinline!array_layout! typedef.
This can be helpful for C++ template metaprogramming.  For example
(C++11 only):
\begin{lstlisting}
template<class ViewType>
void callBlas (const ViewType& A) {
  typedef typename ViewType::array_layout array_layout;
  if (std::is_same<array_layout, LayoutLeft>::value) {
    callSomeBlasFunction (A.ptr_on_device (), ...);
  } else {
    throw std::invalid_argument ("A is not LayoutLeft");
  }
}
\end{lstlisting}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Managing Data Placement}\label{S:Views:Placement}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

TODO: Hey Christian, what did you mean by this?

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Access traits}\label{S:Views:AccessTraits}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Const Views}

Different optional template parameters of View let you control how
Kokkos accesses entries of the array.  The simplest example involves
making a View \lstinline!const!, that is, making its entries read-only.  It
is a compile-time error to assign to an entry of a const View.  It
does not make sense to allocate a const View, but it does make sense
to assign a non-const View to a const View.  Here is an example:
\begin{lstlisting}
const size_t N0 = ...;
View<double*> a_nonconst ("a_nonconst", N0);

// Assign a nonconst View to a const View
View<const double*> a_const = a_nonconst;
// Pass the const View to some read-only function.
const double result = readOnlyFunction (a_const);
\end{lstlisting}
Note that a ``const View'' means the \emph{entries} are const.  You
may still assign to a const View.  \lstinline!View<const double*>!
corresponds exactly to \lstinline!const double*!, and
\lstinline!const View<double*>! to \lstinline!double* const!.

\subsection{Random read Views}

Const Views are useful for declaring intent and helping you debug at
compile time.  They also open the door to other, optional
optimizations.  In particular, Kokkos lets you declare that you plan
to access a read-only array irregularly (less spatial locality), using
the optional \lstinline!RandomAccess! template parameter.  You may only use
this with a const View.  For example:
\begin{lstlisting}
const size_t N0 = ...;
View<int*> a_nonconst ("a", N0); // allocate nonconst View
// Assign to const, RandomAccess View
View<const int*, RandomAccess> a_ra = a_nonconst;
\end{lstlisting}
For a View whose execution space is \lstinline!Cuda!, access to a
\lstinline!RandomAccess! View may use CUDA texture fetches.  Texture
fetches are not cache coherent with respect to writes, which is why
you must use read-only access.  The texture cache is optimized for
noncontiguous access, since it has a shorter cache line than the
regular cache.

\lstinline!RandomAccess! is valid for other execution spaces, and Kokkos
also requires a const View for those execution spaces.  Kokkos
reserves the right to perform analogous optimizations on other
execution spaces.

\subsection{Standard idiom for specifying access traits}

The standard idiom for View is to pass it around using as few template
parameters as possible.  Then, assign to a View with the desired
access traits only at the ``last moment,'' when those access traits
are needed just before entering a computational kernel.  This lets you
template C++ classes and functions on the View type, without
proliferating instantiations.  Here is an example:

\begin{lstlisting}
// Compute a sparse matrix-vector product, for a sparse
// matrix stored in compressed sparse row (CSR) format.
void
spmatvec (const View<double*>& y, 
  const View<const size_t*>& ptr,
  const View<const int*>& ind,
  const View<const double*>& val,
  const View<const double*>& x)
{
  // Access to x has less locality than access to y.
  View<const double*, RandomAccess> x_ra;
  typedef View<const size_t*>::size_type size_type;

  parallel_for (y.dimension_0 (), [=] (const size_type i) {
      double y_i = y(i);
      for (size_t k = ptr(i); k < ptr(i+1); ++k) {
        y_i += val(k) * x_ra(ind(k));
      }
      y(i) = y_i;
    });
}
\end{lstlisting}

\subsection{Unmanaged Views}

It's always better to let Kokkos control memory allocation, but
sometimes you don't have a choice.  You might have to work with an
application or an interface that returns a raw pointer, for example.
Kokkos lets you wrap raw pointers in an \emph{unmanaged View}.
``Unmanaged'' means that Kokkos does \emph{not} do reference counting
or automatic deallocation for those Views.  The following example
shows how to create an unmanaged View of host memory.  You may do this
for CUDA device memory too, or indeed for memory allocated in any
memory space, by specifying the View's execution or memory space
accordingly.
\begin{lstlisting}
// Sometimes other code gives you a raw pointer, ...
const size_t N0 = ...;
double* x_raw = malloc (N0 * sizeof (double));
{
  // ... but you want to access it with Kokkos.
  //
  // malloc() returns host memory, so we use the 
  // host memory space HostSpace.  Unmanaged
  // Views have no label, because labels work with
  // the reference counting system.
  View<double*, HostSpace, Unmanaged> x_view (x_raw, N0);

  functionThatTakesKokkosView (x_view);

  // It's safest for unmanaged Views to fall out of
  // scope, before freeing their memory.
}
free (x_raw);
\end{lstlisting}

\subsection{Atomic access}

The \lstinline!Atomic! access trait lets you create a View, such that every
read or write to any entry uses an atomic update.  This only works for
certain data types, for which the hardware and programming model
support atomic updates.  Currently, Kokkos limits atomic updates to
32-bit or 64-bit built-in types.