### 2015-03-27 Kokkos_PG_LegacyCodeIntegration.tex
Added new chapter which will discuss legacy code transisition and interoperability

\chapter{Interoperability and Legacy Codes}
\section{Legacy data structures}

There are two principal mechanism to facilitate interoperability with legacy data 
structures: (i) Kokkos allocates data and raw pointers are extracted to create 
legacy data structures and (ii) unmanaged views can be used to view externally 
allocated data. In both cases it is mandatory to fix the Layout of the Kokkos view
to the actual layout used in the legacy data structure.

\subsection{Kokkos manages memory}

\subsubsection{Raw allocations}
\subsubsection{Views as the fundamental data owning structure}

\subsection{External memory management}

\subsection{\lstinline|std::vector|}

\section{Calling non-Kokkos libraries}

\section{Keeping track of data transfers}