### 2019-08-29 latex/access_policy_table.tex
talk fully outlined

\begin{tabular}{|c|p{15em}|}
\hline
Expression & Meaning \\
\hline
    \texttt{A} & An \texttt{AccessPolicy} type. \\
    \texttt{a} & An instance of \texttt{A}. \\
    \texttt{A::element\_type} & The type of each element in the set of elements described by the associated \texttt{mdspan}. \\
    \texttt{A::pointer} & The pointer type through which a range of elements are accessed. \\
    \texttt{p} & An instance of \texttt{A::pointer}. \\
    \texttt{i} & Non-negative value of type \texttt{ptrdiff\_t}. \\
    \texttt{A::reference} & The type through which an element is accessed.  Must be convertible to \texttt{A::element\_type} \\
    \texttt{A::offset\_policy} & (Optional) An \texttt{AccessPolicy} type convertible from \texttt{A}. Defaults to \texttt{A}. \\
    \texttt{a.access(p, i)} & Returns an object that provides access to the \texttt{i}-th element in the range of elements that starts at \texttt{p}. \\
    \texttt{a.offset(p, i)} & An instance of \texttt{A::offset\_policy::pointer} for which \texttt{A::offset\_policy(a).access( a.offset(p, i), 0)} references the same element as \texttt{a.access(p, i)}. \\
\hline
\end{tabular}