### 2019-09-01 latex/machine_table.tex
Adding machine table

\begin{tabular}{|l|l|l|}
\hline
Machine | Hardware | Compilers \\
\hline
Apollo & NVIDIA TitanV & CUDA 10.1 (nvcc),GCC 5.3.0 \\
Astra & Cavium Thunder-X2 CN99xx & GCC 8.2 (ARM) \\ 
Blake & Intel(R) Xeon(R) Platinum 8160 & Intel 19.3.199, Intel 18.5.0 \\
Mutrino & Intel(R) Xeon(R) CPU E5-2698 v3 & Intel 18.5.0 \\
\hline
\end{tabular}