### 2020-01-20 src/6.Conclusion.tex

\section{Conclusion}\label{chap:conclusion}
...
Consequently, we emphasize the importance and welcome new development and educational use of parallel programming models that align to this trend. We believe that abstracting from vendor or architecture-specific programming models not only simplifies parallel programming in science and industry but also would support educators in teaching parallel programming.
We would also like to express gratitude to educators for taking on this important role. Further materials, including a detailed description of the programming- and machine model can be found by following the reference on the Kokkos programming mode. Finnaly, we would like to mention that interested developers and student can reach the development team on Slack by following the invititation link as mentioned in the reference section.