### 2020-01-23 src/6.Tutorials.tex
Renamed files for consistency.

\section{Tutorials}\label{chap:tutorial}

Describing here the tutorial app and the example apps that are included either in the repo or in the Kokkos distribution.