### 2020-01-24 main.tex
Added bibitems

\documentclass[conference]{IEEEtran}
\IEEEoverridecommandlockouts
% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{fancyvrb} 
\usepackage{subcaption}
\usepackage{xcolor}
\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
\begin{document}

\title{Performance-portable Parallel Programming in Computer Science Education with Kokkos}

\author{
\IEEEauthorblockN{Jan Ciesko}
\IEEEauthorblockA{\textit{Computer Science Reseach Institute} \\
\textit{Sandia National Laboratories}\\
Albuquerque, NM, USA\\
jciesko@sandia.gov}
\and
\IEEEauthorblockN{David Poliakoff}
\IEEEauthorblockA{\textit{Computer Science Reseach Institute} \\
\textit{Sandia National Laboratories}\\
Albuquerque, NM, USA\\
dzpolia@sandia.gov}
\and
\IEEEauthorblockN{David S. Hollman}
\IEEEauthorblockA{\textit{Computer Science Reseach Institute} \\
\textit{Sandia National Laboratories}\\
Livermore, CA, USA\\
dshallm@sandia.gov}
\and
\IEEEauthorblockN{Author Name}
\IEEEauthorblockA{\textit{Computer Science Reseach Institute} \\
\textit{Sandia National Laboratories}\\
Albuquerque, NM, USA\\
aname@sandia.gov}
\and
\IEEEauthorblockN{Author Name}
\IEEEauthorblockA{\textit{Computer Science Reseach Institute} \\
\textit{Sandia National Laboratories}\\
Albuquerque, NM, USA\\
aname@sandia.gov}
\and
\IEEEauthorblockN{Author Name}
\IEEEauthorblockA{\textit{Computer Science Reseach Institute} \\
\textit{Sandia National Laboratories}\\
Albuquerque, NM, USA\\
aname@sandia.gov}
}

\maketitle

\input{src/0.Abstract}

\begin{IEEEkeywords}
Parallel programming, accelerator programming, Kokkos, C++
\end{IEEEkeywords}

\input{src/1.Introduction}           %label: chap:introduction
\input{src/2.GenericProgramming}     %label: chap:background
\input{src/3.MachineModel}           %label: chap:kokkosPM
\input{src/4.ProgammingModel}        %label: chap:kokkosEMA
\input{src/5.Example}                %label: chap:example
\input{src/6.Tutorials}              %label: chap:tutorial
\input{src/7.C++}                    %label: chap:c++
\input{src/8.Related}                %label: chap:related
\input{src/9.Conclusion}             %label: chap:conclusion

\begin{thebibliography}{00}
\bibitem{RAJA} Hornung, Rich, Jones, Holger, Keasler, Jeff, Neely, Rob, Pearce, Olga, Hammond, Si, Trott, Christian, Lin, Paul, Vaughan, Courtenay, Cook, Jeanine, Hoekstra, Rob, Bergen, Ben, Payne, Josh, and Womeldorff, Geoff., ASC Tri-lab Co-design Level 2 Milestone Report 2015, United States, 2015
\bibitem{KOKKOS} H. Carter Edwards and Christian R. Trott and Daniel Sunderland, 
Kokkos: Enabling manycore performance portability through polymorphic memory access patterns, Journal of Parallel and Distributed Computing, Vol. 74, p. 3202-3216, 2014
\bibitem{NALU} Stefan Domino, Shreyas Ananthan, Robert Knaus, and Alan Williams, 
Deploy Nalu/Kokkos algorithmic infrastructure with performance benchmarking, ECP Milestone Report, United States, 2017
\bibitem{KOKKOSUSECASE1} Holmen, John K. and Humphrey, Alan and Sunderland, Daniel and Berzins, Martin, Improving Uintah’s Scalability Through the Use of Portable Kokkos-Based Data Parallel Tasks, Proceedings of the Practice and Experience in Advanced Research Computing 2017 on Sustainability, ACM, United States, 2017
\bibitem{KOKKOSUSECASE2} Stan Moore, LAMMPS KOKKOS Package: The quest for performance portable MD, LAMMPS Workshop, United States, 2019
\bibitem{OPENMP} https://www.openmp.org/specifications/
\bibitem{ERLANG} Cesarini, Francesco and Thompson, Simon, ERLANG Programming, O’Reilly Media, Inc., 2009
\bibitem{KOKKOS_REPO} https://github.com/kokkos/
\bibitem{KOKKOS_TUTORIAL} https://github.com/kokkos/kokkos-tutorials/
\end{thebibliography}

\end{document}