### 2020-01-24 src/9.Conclusion.tex

\section{Conclusion}\label{chap:conclusion}
%...
In this work we gave an introduction to the Kokkos programming model. We presented a conceptual framework consi,  presented the Kokkos parallel prgroamming model. 
Programming support for parallel progrmaming in mainstream langiages is gaining momentung
. Programming abstraction libraries such as Kokkos, Raja and others showcase this trend. Targeting a generic parallel machine representation allows to design and efficiently use

We emphasize the importance and welcome new development and educational use of parallel programming models that align to this trend. We believe that abstracting from vendor or architecture-specific programming models not only simplifies parallel programming in science and industry but also would support educators in teaching parallel programming.
We would also like to express gratitude to educators for taking on this important role. Further materials, including a detailed description of the programming- and machine model can be found by following the reference on the Kokkos programming mode. Finally, we would like to mention that interested developers and student can reach the development team on Slack by following the invitation link as mentioned in the reference section.