### 2020-01-25 main.tex
improved author formatting.

\documentclass[conference]{IEEEtran}
\IEEEoverridecommandlockouts
% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{fancyvrb} 
\usepackage{subcaption}
\usepackage{xcolor}
\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
\begin{document}

\title{Performance-portable Parallel Programming in Computer Science Education with Kokkos}

\author{
\IEEEauthorblockN{Jan Ciesko}
\IEEEauthorblockA{\textit{Computer Science Reseach Institute} \\
\textit{Sandia National Laboratories}\\
Albuquerque, NM, USA\\
jciesko@sandia.gov}
\and
\IEEEauthorblockN{David Poliakoff}
\IEEEauthorblockA{\textit{Computer Science Reseach Institute} \\
\textit{Sandia National Laboratories}\\
Albuquerque, NM, USA\\
dzpolia@sandia.gov}
\and
\IEEEauthorblockN{David S. Hollman}
\IEEEauthorblockA{\textit{Computer Science Reseach Institute}\\
\textit{Sandia National Laboratories}\\
Livermore, CA, USA\\
dshallm@sandia.gov}
\and
\IEEEauthorblockN{
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ Christian C. Trott }
\IEEEauthorblockA{
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \textit{Computer Science Reseach Institute} \\
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \textit{Sandia National Laboratories}\\
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \  Albuquerque, NM, USA\\
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ ctrott@sandia.gov}
\and
\IEEEauthorblockN{ Damien Lebrun-Grandié\ \ \ \ \ \ \ \ \ \ \ }
\IEEEauthorblockA{\textit{Computational Sciences and Engineering}\ \ \ \ \ \ \ \ \ \ \ \\
\textit{Oak Ridge National Laboratory}\ \ \ \ \ \ \ \ \ \ \ \\
Oak Ridge, TN, USA\ \ \ \ \ \ \ \ \ \ \ \\
lebrungrandt@ornl.gov\ \ \ \ \ \ \ \ \ \ \ }
}

\maketitle

\input{src/0.Abstract}

\begin{IEEEkeywords}
Parallel programming, accelerator programming, Kokkos, C++
\end{IEEEkeywords}

\input{src/1.Introduction}           %label: chap:introduction
\input{src/2.GenericProgramming}     %label: chap:background
\input{src/3.MachineModel}           %label: chap:kokkosPM
\input{src/4.ProgammingModel}        %label: chap:kokkosEMA
\input{src/5.Example}                %label: chap:example
\input{src/6.Backend}                %label: chap:backend
\input{src/8.Related}                %label: chap:related
\input{src/9.Conclusion}             %label: chap:conclusion

\bibliographystyle{IEEEtran}
\bibliography{ref}
\end{document}