### 2020-01-25 src/6.Backend.tex

\section{Back-end Support}\label{chap:backend}

In this section we peak inside the Kokkos library and show how the a generic parallel for loop is mapped to the OpenMP back-end programming model. As a template metaprogramming library, Kokkos makes use of partial specialization. In particular the~\emph{parallel\_for} or ~\emph{parallel\_reduce} functional calls as shown in Figure~\ref{fig:KokkosExample} are mapped to partially specialized classes. Such classes correspond to particular back-end where for each back-end, concepts with partial specialization must be provided. Today, Kokkos supports the CUDA and OpenMP back-ends. For each new back-end, partially implemented classes that implement necessary concepts must be provided.

A simplified version of a partially specialized class for a parallel for loop implementation is shown in Figure\ref{fig:KokkosExampleOMPBackEnd}. The ~\emph{parallel\_for} function instantiates a partial specialization of the ~\emph{ParallelFor} class and calls its ~\emph{execute} function. That function is then responsible to implement the Kokkos parallel pattern. In this back-end, the parallel implementation is generated by using OpenMP pragma annotations and a by invoking a compatible compiler when the library is compiled. 
\begin{figure}
\begin{small}
\begin{minted}[fontsize=\small]{c++}
template <class FunctorType>
class ParallelFor<FunctorType,
              RangePolicy<OpenMP>> 
{
  const FunctorType functor;
  const RangePolicy<OpenMP> policy; 
public:
  void execute() const {
    #pragma omp parallel for
    for(int i=policy.begin();
            i<policy.end();
            ++i)
      functor(i);
  }
};
template <class FunctorType,class PolicyType>
void parallel_for(std::string label, 
    const PolicyType& policy,
    const FunctorType& functor) {
  // Call profiling tool hooks with label
  ParallelFor<FunctorType,PolicyType> 
    pf{functor,policy};
  pf.execute();
}
\end{minted}
\end{small}
\caption{This code shows the back-end implementation of a parallel loop in the Kokkos programing model. It shows a particular implementation of a generic pattern as a templated class to the OpenMP execution space. Also it shows the use of OpenMP pragma annotations that put platform compilers in charge to generate parallel code for the underlying hardware architecture.}
\label{fig:KokkosExampleOMPBackEnd}
\end{figure}

%\begin{figure}
%\centerline{\includegraphics[width=0.3\textwidth]{img/Build.png}}
%\caption{The compilation workflow includes the invocation of the platform compiler that is in charge to generate parallel   Building workflow}
%\label{fig:workflow}  
%\end{figure}