### 2020-02-10 src/0.Abstract.tex
Minor changes. Added pdf output.

\begin{abstract}\label{chap:abstract}
Kokkos is a generic, performance-portable parallel programming model for host- and accelerator architectures in C++. It consists of a pure C++ interface, a specification and a programming library. The programming library exposes patterns and types and maps them to an underlying abstract machine model. This machine model offers a generic view of parallel hardware by abstracting from architecture- and vendor-specific programming models. In academia, this can contribute to a generic eduction of performance-portable parallel programming. In this work we present Kokkos from the perspective of educators, highlight key concepts and point the reader to examples and tutorials suited for incremental teaching of parallel programming in class.
\end{abstract}

%is aligned with C++ standardization efforts and trends towards parallel programming support in base languages.