### 2020-08-10 src/1.Introduction.tex

Parallel and distributed computing is a focus at universities and teaching institutions. The interest is two-fold. Firstly, given their role, educators adapt to trends in research and industry and include parallel computing in their curricula, and secondly, parallel computing is an attractive field for department research and industry collaborations.

While industry-driven collaborations often predefine technological frameworks such as programming environments and targeted architectures, academic teaching leaves this mostly undefined. In this case, educators can select from a variety of solutions, approaches, and trends in parallel computing. Which particular set to choose for teaching in class is not trivial. Complexity is further amplified by the breath of related topics ranging from parallel hardware architectures, programming models, algorithms as well as productivity in parallel software development. However, it is the productivity concern that is gaining importance as the number of hardware solutions, their complexity as well as demands for a quick time-to-solution increase.

While vendor programming models maintain their position in curricula today, we think that the inclusion of parallel programming models such as Kokkos is worth consideration. They offer an easy to understand set of concepts for a vendor- and architecture-independent, performance-portable expression of concurrency which significantly contributes to parallel programming productivity.

Kokkos\footnote{from Greek for grain or seed}~\cite{KOKKOS} is a performance-portable, parallel programming model for C++ developed collaboratively by several national laboratories as part of the US Department of Energy Exascale Project\cite{ECP}. It consists of a programming model specification and a library. The specification defines the execution and memory model and an application programming interface (library API). The runtime library exposes programming primitives, namely execution patterns and types, including views and execution spaces and maps them to the underlying abstract machine model. The machine model creates a generic view of hardware and memories. Finally, the Kokkos library implements the mapping of the programming model to vendor-specific programming models, libraries, and memory layouts. Kokkos has been successfully used in many large-scale projects at Sandia National Laboratories, Oak-Ridge National Laboratories, Los Alamos National Laboratories, University of Utah, and others~\cite{KOKKOSPROJS}. We list its key properties as follows.

\begin{itemize}
\item Targets an abstract machine model
\item Implements patterns, views, and spaces
\item Vendor- and architecture-independent and open-source
\item Aligns with C++ standardization efforts for parallel programming
\item Includes exercises and teaching material
\end{itemize}

The objective of this paper is to present Kokkos from three perspectives. Firstly, we would like to present our thoughts that accompanied the design and development of the Kokkos parallel programming model. These insights may be useful in parallel programming classes that introduce programming models, reason about design choices, and cover productivity in parallel software development.

Further, we would like to promote Kokkos to students and educators as a platform that helps them to understand and experiment with parallel patterns in a way that is generic, vendor-agnostic, and declarative. Abstractions allow students to learn about concurrency without being exposed to many architectural concerns initially and help to create awareness of their productivity benefits in terms of code- and performance portability on current and upcoming hardware architectures.

Finally, we would like to present Kokkos to the educational community as one representative parallel programming model for a trend towards portability abstractions in parallel programming models. We show a set of interfaces that implement patterns, views, and spaces, and capture enough semantic information to produce performance-portable, parallel code on current architectures. Further, we highlight its alignment and contributions to standardization efforts towards supporting parallel programming in C++. For teaching insights, we document feedback received during on-line lectures.

The rest of this paper is structured as follows. The next chapter gives an overview of parallel programming concepts and presents the notion of \emph{semantic capture}. Chapter~\ref{chap:kokkosMM} presents the abstract machine model used to derive the required semantic information for performance portability in Kokkos. Chapters~\ref{chap:kokkosPM} and ~\ref{chap:kokkosExample} present the Kokkos programming model and an example application. In Chapter~\ref{chap:backend} we provide insight into a Kokkos back-end implementation. Chapters~\ref{chap:related} and ~\ref{chap:tutorials} discuss related work, tutorials and student feedback from tutorials. Lastly Chapter~\ref{chap:conclusion} concludes this work.